Recovery of deleted deep sequencing data sheds more light on the early Wuhan SARS-CoV-2 epidemic

### 2021-06-18 paper/paper.tex
few more minor text tweaks

\lettrine[lines=2]{\color{color2}U}{}nderstanding the spread of SARS-CoV-2 in Wuhan is crucial to tracing the origins of the virus, including identifying events that led to infection of patient zero.
The first reports outside of China at the end of December 2019 emphasized the role of the Huanan Seafood Market~\citep{ProMED2109}, which was initially suggested as a site of zoonosis.
However, this theory became increasingly tenuous as it was learned that many early cases had no connection to the market~\citep{cohen2020wuhan,huang2020clinical, chen2020epidemiological}.
Soon, Chinese CDC Director Gao Fu dismissed the theory, stating ``At first, we assumed the seafood market might have the virus, but now the market is more like a victim.
The novel coronavirus had existed long before''~\citep{GlobalTimes2020}.

@@

However, the Wuhan sequences that are available have also confounded phylogenetic analyses designed to infer the ``progenitor'' of SARS-CoV-2~\citep{kumar2021evolutionary}.
Although there is debate about exactly how SARS-CoV-2 entered the human population, it is universally accepted that the virus's deep ancestors are bat coronaviruses~\citep{lytras2021exploring}.
But the earliest known SARS-CoV-2 sequences, which are mostly derived from the Huanan Seafood Market, are notably more different from these bat coronaviruses than other SARS-CoV-2 sequences collected at later dates outside Wuhan.
As a result, there is a direct conflict between the two major principles used to infer an outbreak's progenitor: namely that it should be among the earliest sequences, and that it should be most closely related to deeper ancestors~\citep{pipes2021assessing}.

@@

\section{Competing interests}
The author consults for Moderna on SARS-CoV-2 evolution and epidemiology, consults for Flagship Labs 77 on viral deep mutational scanning, and has the potential to receive a share of IP revenue as an inventor on a Fred Hutch licensed technology/patent (application WO2020006494) related to deep mutational scanning of viral proteins.

>>>>

\lettrine[lines=2]{\color{color2}U}{}nderstanding the spread of SARS-CoV-2 in Wuhan is crucial to tracing the origins of the virus, including identifying events that led to infection of patient zero.
The first reports outside of China at the end of December 2019 emphasized the role of the Huanan Seafood Market~\citep{ProMED2109}, which was initially suggested as a site of zoonosis.
However, this theory became increasingly tenuous as it was learned that many early cases had no connection to the market~\citep{cohen2020wuhan,huang2020clinical, chen2020epidemiological}.
Eventually, Chinese CDC Director Gao Fu dismissed the theory, stating ``At first, we assumed the seafood market might have the virus, but now the market is more like a victim.
The novel coronavirus had existed long before''~\citep{GlobalTimes2020}.

@@

However, the Wuhan sequences that are available have also confounded phylogenetic analyses designed to infer the ``progenitor'' of SARS-CoV-2, which is the sequence from which all other currently known sequences are descended~\citep{kumar2021evolutionary}.
Although there is debate about exactly how SARS-CoV-2 entered the human population, it is universally accepted that the virus's deep ancestors are bat coronaviruses~\citep{lytras2021exploring}.
But the earliest known SARS-CoV-2 sequences, which are mostly derived from the Huanan Seafood Market, are notably more different from these bat coronaviruses than other sequences collected at later dates outside Wuhan.
As a result, there is a direct conflict between the two major principles used to infer an outbreak's progenitor: namely that it should be among the earliest sequences, and that it should be most closely related to deeper ancestors~\citep{pipes2021assessing}.

@@

\section{Competing interests}
The author consults for Moderna on SARS-CoV-2 evolution and epidemiology, consults for Flagship Labs 77 on viral evolution and deep mutational scanning, and has the potential to receive a share of IP revenue as an inventor on a Fred Hutch licensed technology/patent (application WO2020006494) related to deep mutational scanning of viral proteins.
